### 2021-07-06 Draft.tex
abstract finished

%\documentclass[12pt, a4paper]{article}
\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{ngerman}
\usepackage[ansinew]{inputenc}
\usepackage[left=2.5cm,top=2.5cm]{geometry}

\renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}

\title{The effect of commen network problems on students academic performance in an elearning-Environment \thanks{Supported by Goethe University Frankfurt a. Main}}

\author{Lucas Laub\inst{1}\orcidID{6621331} \and
Second Author\inst{2}\orcidID{1111-2222-3333-4444}}

\institute{Goethe University Frankfurt, 60323 Frankfurt a. Main, Germany.}

\maketitle

\begin{abstract}
In the current light of the pandemic the worldwide use
of eLearning-Software experienced an unpresented boom.
We state the question how commen network problems influence
the academic performance in an eLearning-Environment.
To provide answers an online questionnaire with deliberate
technical difficulties was constructed. Evaluating the performance
of the test and controll group did not show any significant
differences.
\keywords{eLearning  \and Online-Learning \and academic performance.}
\end{abstract}

\section{Introduction}
\section{Materials and Methodes}
\section{Results}
\section{Discussion}
\section{Conclusion}
\section{Refernces}
\end{document}